\bf Landslide: \\ Systematic Exploration for Kernel-Space Race Detection

### 2012-06-05 rv/paper.tex
Write introduction

%%%%%%%%%%%%%%%%%%%%%%%%
\title{\bf Landslide: Systematic Exploration for Kernel-Space Race Detection}
\author{Ben Blum \and Garth Gibson}
\institute{Computer Science Department, Carnegie Mellon University, Pittsburgh}
\date{May 2012}

@@

@@

% put this somewhere
%We claim that this combination of the user's knowledge and Landslide's testing mechanisms leads to an effective usage dynamic, in which the user steers Landslide towards focused search spaces that are more likely to find bugs.

@@

\revision{
Many techniques exist for dynamic testing of concurrent systems for race conditions.
Systematic exploration~\cite{verisoft}, the strategy we focus on in this work, involves making educated guesses as to what points during execution a preemption would be most likely to expose a bug, enumerating the different possibilities for interleaving threads around these points, and forcing the system to execute all such interleavings to check if any of them results in incorrect behaviour.
Systematic exploration provides a better alternative to conventional long-running stress tests, because it is less likely to overlook buggy execution patterns, and it enables a testing framework to report more thorough debugging information. Compared to other dynamic analyses, such as data race detection~\cite{datacollider}, systematic exploration is able to find a wider range of types of concurrency errors because of its ability to manipulate the execution of the system under test.

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested. While user-space testing frameworks rely on the underlying kernel to provide common concurrency abstractions which the framework can use to drive the test, systematic exploration in the kernel itself necessarily exists underneath such abstractions.
Furthermore, a tool that supports testing multiple kernels must support many different possible scheduler designs.

The claim of this dissertation is as follows.
\begin{quote}
\bf
Systematic exploration in kernel space is possible by understanding a kernel's concurrency through instrumentation of its internal abstractions, and can be made efficient by relying on the user to control the scope of the test.
\end{quote}
}

We present Landslide,\footnote{
Landslide {\em (n)}: a phenomenon which demonstrates that Pebbles are not as stable as you might think.}
a tool for applying systematic race detection techniques to kernel-level code.
It is geared towards kernels that implement the Pebbles specification, the main project in Operating System Design and Implementation (15-410) at CMU, and is implemented as a module for Wind River Simics\textsuperscript{\texttrademark}~\cite{simics}, the x86 simulator that students use to run their Pebbles kernels.
\revision{In order to use Landslide, a kernel programmer must instrument their kernel to inform Landslide of important concurrency events during the execution, and configure Landslide to focus its search on components of the kernel most relevant to the test.
%During a kernel's execution, Landslide records important actions performed by the kernel, decides at which points in the kernel's execution a preemption will be most likely to expose a bug, and then exercises all possible interleavings of kernel threads around such points.
Using this instrumentation and configuration, Landslide is able to enumerate all possible interleavings of kernel threads at a granularity both fine-grained enough to expose interesting concurrency behaviour and coarse-grained enough to produce state spaces that are feasible to explore. Landslide then forces the kernel to exercise each of these interleavings, and checks for race conditions in each of them.}
When Landslide finds a bug (predicted by a set of common checks and heuristics), it stops execution and prints information about the sequence of interleavings that exposed the bug.

\revision{
In this work, we study the challenges inherent in applying systematic testing techniques in kernel-space, in contrast with user-space applications (Chapter~\ref{sec:challenges}), define common terms used in the rest of this document (Chapter~\ref{sec:terminology}), present our techniques (some sound, and some heuristic) for addressing them (Chapter~\ref{sec:design}), and document the process of using Landslide (Chapter~\ref{sec:using}).
We evaluate Landslide both by studying its potential to be a helpful debugging aid in 15-410 and by studying certain bugs in detail to determine effective usage strategies (Chapter~\ref{sec:evaluation}).
Finally, we discuss possible avenues for future work in education, use of additional testing techniques, application to general purpose kernels, long-running testing approaches, and theoretical study of systematic exploration (Chapter~\ref{sec:future}).
}

%\input{related} % COMPLETE!
%\input{challenges} % COMPLETE!
%\input{terminology} % almost complete
%\input{design}
%\input{using}
%\input{evaluation}
%\input{futurework}
%\input{conclusion}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
\title{\bf Landslide: \\ Systematic Exploration for Kernel-Space Race Detection}
\author{Ben Blum \and Garth Gibson}
\institute{Computer Science Department, Carnegie Mellon University, Pittsburgh}
\date{May 2012}

@@

%%%%% better outline %%%%
% intro
% related
% challenges of kernel-space (short section, 2-3 paragraphs)
% - causes of concurrency
% - ad-hoc memory sharing
% scope and model
% - summary of 410 & kernel requirements
% - simulated execution
% - timer-driven scheduling/concurrency
% - orientation for false-negative bugs & coarse-granularity search
% implementation
% - components (very briefly; only keep challenge-solving stuff)
% - bug identification techniques
% - making DPOR work
% user interface
% - instrumenting process
% - iterative configuring + focusing of search
% - types of test cases to use
% future work
% - new techniques to incorporate
% - virtualisation and linux
% - maybe shaping? (not sure how i could say it quickly.)

@@

@@

Systematic exploration~\cite{verisoft} is a dynamic testing technique that aims to find race conditions or verify their absence. It involves enumerating many different interleavings of multiple threads and forcing a system to execute all such interleavings to check if any of them expose bugs.
In comparison with conventional stress testing, systematic exploration is less likely to overlook bugs due to its deterministic testing approach. While other dynamic analysis techniques exist for finding race conditions, such as data race detection, % TODO cite here
systematic exploration is able to find a wider range of types of concurrency errors because of its ability to manipulate the execution of the system under test.

%\revision{
%Many techniques exist for dynamic testing of concurrent systems for race conditions.
%Systematic exploration~\cite{verisoft}, the strategy we focus on in this work, involves making educated guesses as to what points during execution a preemption would be most likely to expose a bug, enumerating the different possibilities for interleaving threads around these points, and forcing the system to execute all such interleavings to check if any of them results in incorrect behaviour.
%Systematic exploration provides a better alternative to conventional long-running stress tests, because it is less likely to overlook buggy execution patterns, and it enables a testing framework to report more thorough debugging information. Compared to other dynamic analyses, such as data race detection~\cite{datacollider}, systematic exploration is able to find a wider range of types of concurrency errors because of its ability to manipulate the execution of the system under test.

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested.
While user-space testing frameworks operate above such abstractions, a systematic exporation tool for kernels cannot rely on them.
Furthermore, execution in the kernel's scheduler and other similar components is common to all concurrent threads, which threatens our ability to achieve state-space reduction.

%While user-space testing frameworks rely on the underlying kernel to provide common concurrency abstractions which the framework can use to drive the test, systematic exploration in the kernel itself necessarily exists underneath such abstractions.
%Furthermore, a tool that supports testing multiple kernels must support many different possible scheduler designs.

% The claim of this dissertation is as follows.
% \begin{quote}
% \bf
% Systematic exploration in kernel space is possible by understanding a kernel's concurrency through instrumentation of its internal abstractions, and can be made efficient by relying on the user to control the scope of the test.
% \end{quote}
% }

We present Landslide,
%\footnote{Landslide {\em (n)}: a phenomenon which demonstrates that Pebbles are not as stable as you might think.}
a tool for applying systematic race detection techniques to kernel-level code.
Landslide is implemented as a module for Wind River Simics\textsuperscript{\texttrademark}~\cite{simics}, and targets Pebbles, a fully preemptible UNIX-like kernel specification that students implement in Carnegie Mellon University's undergraduate operating systems class (15-410).
To use Landslide, a programmer must instrument their kernel to inform Landslide of important concurrency events during the execution. In order to make Landslide's search efficient, the programmer should also configure Landslide to focus on the most relevant parts of the kernel.
This leads to a usage dynamic that combines the user's design knowledge and Landslide's exploration mechanisms. Intuitively, the user ``steers'' Landslide towards focused search spaces that are more likely to quickly find bugs.

%It is geared towards kernels that implement the Pebbles specification, the main project in Operating System Design and Implementation (15-410) at CMU, and is implemented as a module for Wind River Simics\textsuperscript{\texttrademark}~\cite{simics}, the x86 simulator that students use to run their Pebbles kernels.
%\revision{In order to use Landslide, a kernel programmer must instrument their kernel to inform Landslide of important concurrency events during the execution, and configure Landslide to focus its search on components of the kernel most relevant to the test.
%During a kernel's execution, Landslide records important actions performed by the kernel, decides at which points in the kernel's execution a preemption will be most likely to expose a bug, and then exercises all possible interleavings of kernel threads around such points.
%Using this instrumentation and configuration, Landslide is able to enumerate all possible interleavings of kernel threads at a granularity both fine-grained enough to expose interesting concurrency behaviour and coarse-grained enough to produce state spaces that are feasible to explore. Landslide then forces the kernel to exercise each of these interleavings, and checks for race conditions in each of them.}
%When Landslide finds a bug (predicted by a set of common checks and heuristics), it stops execution and prints information about the sequence of interleavings that exposed the bug.

We found that, with modest instrumentation effort, Landslide was able to find race conditions in several different Pebbles kernels that had previously been overlooked by stress testing and/or expert manual inspection.
Many of these bugs required complicated thread interleavings to expose, and could not have been found with simpler testing techniques such as data race detection or static analysis.

%\revision{
%In this work, we study the challenges inherent in applying systematic testing techniques in kernel-space, in contrast with user-space applications (Chapter~\ref{sec:challenges}), define common terms used in the rest of this document (Chapter~\ref{sec:terminology}), present our techniques (some sound, and some heuristic) for addressing them (Chapter~\ref{sec:design}), and document the process of using Landslide (Chapter~\ref{sec:using}).
%We evaluate Landslide both by studying its potential to be a helpful debugging aid in 15-410 and by studying certain bugs in detail to determine effective usage strategies (Chapter~\ref{sec:evaluation}).
%Finally, we discuss possible avenues for future work in education, use of additional testing techniques, application to general purpose kernels, long-running testing approaches, and theoretical study of systematic exploration (Chapter~\ref{sec:future}).
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Challenges of Kernel Space}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scope}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Environment}
% 410, requirements, simulated
\subsection{Testing Approach}
% timer driven, false negative bugs, coarse granularity search

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Design}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% solving challenges %%%%
% timer-driven scheduling
% tracking which threads

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{User Interface}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% decision trace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%We evaluated Landslide in two ways: first, by instrumenting two kernels the authors were familiar with and experimenting with many different configurations for Landslide's search, and second, by meeting with students

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
